---
title: "pam_ed50_relativefvfm_bygeno"
author: "Melissa Naugle"
date: "5/20/2022"
output: html_document
---

# Script to consolidate and quality control sampling metadata, heat toleance phenotypes, and site data
- Remove ED50s with large confidence intervals
- Remove data missign control values
- Generate heat tolerance trait values adjusted for site 

```{r setup, include=FALSE}
setwd("/Users/Melissa/Desktop/GitHub/Ahya_heat_tol_variation/Rscripts_dataprocessing/")
rm( list = ls())
graphics.off()
library(tidyverse)
library(reshape2)
library(gridExtra)
library(PupillometryR)
library(Rmisc)
date = format(Sys.Date(), "%Y_%m_%d")
```

```{r}
chl <- read.csv("../Rawdata/TotalChlorophyll.csv")
chl$Site_geno_year <- paste0(chl$Site,"_",chl$Genotype,"_",chl$SampleYear)
events <- read.csv("../Rawdata/GenomicdatafromtheECT1_RRAP_2023_10_29_depths2023-10-30.csv") %>% dplyr::select(eventID, Site_geno_year)
chl2 <- left_join(chl, events)
write.csv(chl2, "../Rawdata/TotalChlorophyll.csv", row.names = F)
```


```{r }
metadata <- read.csv("../Rawdata/GenomicdatafromtheECT1_RRAP_2023_10_29_depths2023-10-30.csv")

#load fvfm metrics
absoluteed50s <- read.csv("../CleanData/iPAM/ed50s_genos_2023_12_06.csv") 
relativefvfm <- read.csv("../CleanData/iPAM/relativeFvFm_correctedtemps_2023_12_06.csv")

site_characteristics <- read.csv("../Rawdata/Site_characteristics.csv") %>% dplyr::select(-MMM.)

absoluteed50s$Site_geno_year <- paste0(absoluteed50s$Site,"_",absoluteed50s$Genotype,"_",absoluteed50s$SampleYear)

ed50_dat <- left_join(relativefvfm, absoluteed50s,by = "Site_geno_year", suffix = c("", ".z")) %>%
  dplyr::select(-ends_with(".z")) %>%
  distinct()


ipam_data <- left_join(ed50_dat, metadata,by = "Site_geno_year", suffix = c("", ".z")) %>%
  dplyr::select(-ends_with(".z")) %>%
  left_join(. ,site_characteristics,by = "Site.name", suffix = c("", ".z")) %>%
  dplyr::select(-ends_with(".z")) %>%
  distinct() %>%
  dplyr::rename(fvfm_ED50 = Absolute_estimate_ED50) %>%
  dplyr::rename(fvfm_ed50_std_error = Std..Error)%>%
  dplyr::select(-X)

#load hyperspec metrics
ndvi_absoluteed50s <- read.csv("../CleanData/NDVI/NDVI_ed50s_genos_2023_12_06.csv") 
ndvi_relativefvfm <- read.csv("../CleanData/NDVI/relativeNDVI_correctedtemps_2023_12_06.csv")

hyperspecdat <- left_join(ndvi_relativefvfm,ndvi_absoluteed50s, by = "Site_geno_year", suffix = c("", ".z")) %>%
  dplyr::select(-ends_with(".z"))%>%
  dplyr::rename(ndvi_ED50 = Absolute_estimate_ED50) %>%
  dplyr::rename(ndvi_ed50_std_error = Std..Error) %>%
  dplyr::select(-X)

all_data <- left_join(ipam_data,hyperspecdat, by = "Site_geno_year", suffix = c("", ".z")) %>%
  dplyr::select(-ends_with(".z"))

```

#fix health chart and field ID
```{r, include=FALSE}
all_data$Health.Chart.num <- all_data$Health.Chart
all_data$Health.Chart.num <- gsub("D3/4","3.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("E3/4","3.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("C3/4","3.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("C2/3","2.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("D2/3","2.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("E2/3","2.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("C4/5","4.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("D4/5","4.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("E4/5","4.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("2/3","2.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("3/4","3.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("4/5","4.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("5/6","5.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("E2/E3","2.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("E3/E4","3.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("E4/E5","4.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("C2/C3","2.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("C3/C4","3.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("C4/C5","4.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("C5/C6","5.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("D2/D3","2.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("D3/D4","3.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("D4/D5","4.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- gsub("D5/D6","5.5",all_data$Health.Chart.num)
all_data$Health.Chart.num <- parse_number(all_data$Health.Chart.num)

#manually add in missing Health chart data 
all_data[is.na(all_data$Health.Chart.num),]
all_data$Health.Chart.num[all_data$Site_geno_year == "LadyMusgrave_107_2021"] <- 4
all_data$Health.Chart.num[all_data$Site_geno_year == "Sykes_126_2021"] <- 4.5
all_data$Health.Chart.num[all_data$Site_geno_year == "Sykes_127_2021"] <- 5
all_data$Health.Chart.num[all_data$Site_geno_year == "Sykes_128_2021"] <- 5
all_data$Health.Chart.num[all_data$Site_geno_year == "Sykes_129_2021"] <- 3
all_data$Health.Chart.num[all_data$Site_geno_year == "Sykes_130_2021"] <- 5.5
all_data$Health.Chart.num[all_data$Site_geno_year == "Sykes_131_2021"] <- 4.5
all_data$Health.Chart.num[all_data$Site_geno_year == "Sykes_124_2021"] <- 4



#select only relevant variables
all_data <- all_data %>% dplyr::select("Site_year", "Collection.date","Site","Site_geno_year","Site.name","Genotype","Depth_corrected", "Time","Health.Chart","Health.Chart.num","SampleYear","Tide_adjust","Depth_corrected","fvfm_ED50","fvfm_ED50CIrange","LowerCI_fvfmED50","UpperCI_fvfmED50","fvfm_ed50_std_error","ndvi_ED50","ndvi_ED50CIrange","LowerCI_ndviED50","UpperCI_ndviED50","ndvi_ed50_std_error","RelativeNDVI9C","RelativeFvFm9C","Lat","Lon","Km_to_coastline", "Subregion","Shelf.pos","Region","acuteHeatStress" ,"naturalBleaching", "resampled" ,"subproject")

```


Add trip variable
```{r}
#add trip num variable
all_data$Trip <- 0
all_data$Trip[all_data$Site_year == "Reef21550-20Jan_2021" | all_data$Site_year == "Reef21550-21Jan_2021" | all_data$Site_year == "EastCay-22Jan_2021" | all_data$Site_year == "EastCay-23Jan_2021" | all_data$Site_year == "LadyMusgrave_2021" | all_data$Site_year == "Sykes_2021"] <- 1
all_data$Trip[all_data$Site_year == "Hicks-21March_2021" | all_data$Site_year == "Hicks-20March_2021" | all_data$Site_year == "NorthDirection-23March_2021" | all_data$Site_year == "NorthDirection-24March_2021" | all_data$Site_year == "NoName_2021" | all_data$Site_year == "Martin_2021" |all_data$Site_year == "StCrispin-30Mar_2021" ] <- 2

all_data$Trip[all_data$Site_year == "Kelso_2021" | all_data$Site_year == "Pelorus_2021" | all_data$Site_year == "StCrispin-15April_2021" | all_data$Site_year == "Mackay_2021" | all_data$Site_year == "Chicken_2021" | all_data$Site_year == "Davies_2021"] <- 3

all_data$Trip[all_data$Site_year == "Fitzroy_2021" | all_data$Site_year == "Moore_2021" ] <- 4

all_data$Trip[all_data$Site_year == "Mackay_2022" | all_data$Site_year == "Chicken_2022" | all_data$Site_year == "Davies_2022" & all_data$Collection.date == "2022-03-11"] <- 5

all_data$Trip[all_data$Site_year == "LadyMusgrave_2022" | all_data$Site_year == "Heron_2022"] <- 6

```


#quality control 
```{r }
pheno2 <- all_data %>% 
  filter(!Site == "Reef21550-21Jan") %>% #remove Reef21550 run on 21 jan bc bad temp profile did not reach +9
  filter(!Site_geno_year == "Chicken_310_2021") #colony lost and no pheno data collected

#missing ndvi?
pheno2 %>% filter(is.na(ndvi_ED50))

#remove hicks-20 march from NDVI bc MMM is missing
pheno2$ndvi_ED50[pheno2$Site == "Hicks-20March"] <- NA
pheno2$ndvi_ED50CIrange[pheno2$Site == "Hicks-20March"] <- NA
pheno2$LowerCI_ndviED50[pheno2$Site == "Hicks-20March"] <- NA
pheno2$UpperCI_ndviED50[pheno2$Site == "Hicks-20March"] <- NA
pheno2$ndvi_ed50_std_error[pheno2$Site == "Hicks-20March"] <- NA
pheno2$RelativeNDVI9C[pheno2$Site == "Hicks-20March"] <- NA

pheno2 %>% filter(ndvi_ED50CIrange < 5) #490
pheno2 %>% filter(!is.na(ndvi_ED50)) #534
493/534

pheno2 %>% filter(fvfm_ED50CIrange < 5) #476
pheno2 %>% filter(!is.na(fvfm_ED50)) #569
474/569

pheno2[is.na(pheno2$fvfm_ED50),] # 0
pheno2[is.na(pheno2$ndvi_ED50),] # 35
pheno2$ndvi_ED50[pheno2$ndvi_ED50CIrange > 5] <- NA
pheno2$ndvi_ED50[is.na(pheno2$ndvi_ED50CIrange)] <- NA
pheno2$fvfm_ED50[pheno2$fvfm_ED50CIrange > 5] <- NA
pheno2$fvfm_ED50[is.na(pheno2$fvfm_ED50CIrange)] <- NA
pheno2[is.na(pheno2$fvfm_ED50),] # 95 missing
pheno2[!is.na(pheno2$fvfm_ED50),] # 474 total
pheno2[is.na(pheno2$ndvi_ED50),] # 76 missing
pheno2[!is.na(pheno2$ndvi_ED50),] #  493 total
```

# Adjust phenotype values by site 
```{r }

#get values adjusted for site
dat_all_red <- pheno2 %>% filter(!is.na(RelativeFvFm9C))
dat_all_mod <- lm(RelativeFvFm9C ~ Site.name, data = dat_all_red)
dat_all_relfvfm.adj <- broom::augment(dat_all_mod, data = dat_all_red) %>%
  mutate(RelativeFvFm9C.adj = mean(dat_all_red$RelativeFvFm9C) + .resid)%>% dplyr::select(Site_geno_year,RelativeFvFm9C.adj)



dat_all_red <- pheno2 %>% filter(!is.na(RelativeNDVI9C))
dat_all_mod <- lm(RelativeNDVI9C ~ Site.name, data = dat_all_red)
dat_all_relndvi.adj <- broom::augment(dat_all_mod, data = dat_all_red) %>%
  mutate(RelativeNDVI9C.adj = mean(dat_all_red$RelativeNDVI9C) + .resid)%>% dplyr::select(Site_geno_year,RelativeNDVI9C.adj)


dat_all_red <- pheno2 %>% filter(!is.na(ndvi_ED50))
dat_all_mod <- lm(ndvi_ED50 ~ Site.name, data = dat_all_red)
dat_all_ndvi_ED50.adj <- broom::augment(dat_all_mod, data = dat_all_red) %>%
  mutate(ndvi_ED50.adj = mean(dat_all_red$ndvi_ED50) + .resid)%>% dplyr::select(Site_geno_year,ndvi_ED50.adj)


dat_all_red <- pheno2 %>% filter(!is.na(fvfm_ED50))
dat_all_mod <- lm(fvfm_ED50 ~ Site.name, data = dat_all_red)
dat_all_fvfm_ED50.adj <- broom::augment(dat_all_mod, data = dat_all_red) %>%
  mutate(fvfm_ED50.adj = mean(dat_all_red$fvfm_ED50) + .resid) %>% dplyr::select(Site_geno_year,fvfm_ED50.adj)


dat_all.adj <- full_join(pheno2, dat_all_relfvfm.adj, by = "Site_geno_year") %>% full_join(dat_all_relndvi.adj, by = "Site_geno_year")%>% full_join(dat_all_ndvi_ED50.adj, by = "Site_geno_year") %>% full_join(dat_all_fvfm_ED50.adj, by = "Site_geno_year")

```

#make sitename_geno_year variable
```{r}
dat_all.adj$Sitename_geno_year <- paste0(dat_all.adj$Site.name,"_",dat_all.adj$Genotype,"_",dat_all.adj$SampleYear)
```


```{r }
write.csv(dat_all.adj,paste0("../CleanData/pheno_metadata_clean_",date,".csv"), row.names = F)
```





