---
title: "BRT"
author: "Melissa Naugle"
date: "1/20/2023"
output: html_document
---

# Script to build individual and site level BRTs to test effects of predictors on heat tolerance traits
- Select top predictors and plot their effect on traits (Fig 6)

```{r setup, include=FALSE}
setwd("/Users/Melissa/Desktop/GitHub/Ahya_heat_tol_variation//Rscripts_analysis/")
#rm(list = ls())
graphics.off()
library(tidyverse)
library(gbm)
library(dismo)
library(patchwork)
date = format(Sys.Date(), "%Y_%m_%d")
outputpath <- "/Users/Melissa/Desktop/GitHub/Ahya_heat_tol_variation//Outputs/BRTs/"
```

#data in
```{r}
pheno <- read.csv("../CleanData/pheno_metadata_clean_2024_01_08.csv")
env <- read.csv("../CleanData/Environmental_data_PCs_2024_01_08.csv")
clust <- read.csv("../CleanData/2023_10_17_clust_data_edited.csv")
symbiont <- read.csv("../CleanData/ITS2_types_OTUs_17_10_2023.csv") %>% dplyr::select(Site.name_geno_year,dominant_type,all_typeprofs,OTU_PC1, OTU_PC2, OTU_PC3)
pred_types <- read.csv("../CleanData/EnvMetric_categories.csv")

dat_all <- left_join(pheno, env, by = "Site_geno_year",suffix = c("", ".z")) %>%
  dplyr::select(-ends_with(".z"))

dat_all$Sitename_geno_year = paste0(dat_all$Site.name,"_",dat_all$Genotype,"_",dat_all$SampleYear)

dat_all2 <- left_join(dat_all, clust, by = "Sitename_geno_year",suffix = c("", ".z")) %>%
  dplyr::select(-ends_with(".z"))

dat_all2$Site.name_geno_year <- dat_all2$Sitename_geno_year

dat_all3 <- dat_all2 %>% left_join(symbiont, by = "Site.name_geno_year",suffix = c("", ".z")) %>%
  dplyr::select(-ends_with(".z"))

```


# select predictors and set up datasets 
```{r }
#move single occurance dominant types to group into "other" and reduce bias of categorical predictors with too many categories 
dat_all3$dominant_type.adj <- dat_all3$dominant_type
dat_all3 %>% dplyr::select(Site_geno_year,dominant_type.adj) %>% group_by(dominant_type.adj) %>% tally()

dat_all3$dominant_type.adj[dat_all3$dominant_type.adj == "C3k-C50a-C3wy-C50b-C3ba-C3wu"] <- "Other"
dat_all3$dominant_type.adj[dat_all3$dominant_type.adj == "C50b/C50p-C3-C50f-C3bm-C3dt"] <- "Other"

dat_all3 %>% dplyr::select(Site_geno_year,dominant_type.adj) %>% group_by(dominant_type.adj) %>% tally()
```
#Choose predictors to include and set up data to run in BRT models 
```{r }
dat_all3$Region <- "M"
dat_all3$Region[dat_all3$Subregion == "Swains" | dat_all3$Subregion == "Capricorn Bunker"] <- "South"
dat_all3$Region[dat_all3$Subregion == "Cooktown" | dat_all3$Subregion == "Innisfail/Townsville" | dat_all3$Subregion == "Townsville" | dat_all3$Subregion == "Cairns"] <- "North-Central"

#choose factors to include
dat_all_touse <- dat_all3 %>% filter(!is.na(fvfm_ED50)) %>% dplyr::select(c(
"Site_geno_year", 
#temp
"MMM","Depth_corrected","DHW_collection","TSA_DHW_mean",
"TSA_DHW_stdev", "DHW_freq_sup4","DHW_freq_sup8","DHW_max","SSTA_freq_stdev",
"SSTA_freq_mean", "ROTC_ss", "AR" ,"DTR_ss","DTR","OM","Lat",
#nutrients
"TOTAL_NITROGEN_OM","NO3_OM","PIP_OM",
#light
"Secchi_OM","CF_OM",
#symbiont
"dominant_type.adj",
"OTU_PC1","OTU_PC2",
#clust
"Clust",
"Shelf.pos",
#pheno
"fvfm_ED50", "ndvi_ED50","RelativeNDVI9C","RelativeFvFm9C","fvfm_ED50CIrange","ndvi_ED50CIrange" , 
"Region", "Site.name", "Subregion","Health.Chart.num","Site"
))

dat_all_touse$dominant_type.adj <- as.factor(dat_all_touse$dominant_type.adj)
dat_all_touse$Clust <- as.factor(dat_all_touse$Clust)
dat_all_touse$Site <- as.factor(dat_all_touse$Site)
dat_all_touse$Shelf.pos <- as.factor(dat_all_touse$Shelf.pos)



###
dat_all_fvfmED50_filt <- dat_all_touse %>% dplyr::select(-c("ndvi_ED50","RelativeNDVI9C", "RelativeFvFm9C","fvfm_ED50CIrange","ndvi_ED50CIrange","Site_geno_year","Region","Site.name", "Subregion","Site"))%>% filter(!is.na(fvfm_ED50)) 

dat_all_ndviED50_filt <- dat_all_touse %>% dplyr::select(-c("fvfm_ED50","RelativeNDVI9C", "RelativeFvFm9C","fvfm_ED50CIrange","ndvi_ED50CIrange","Site_geno_year","Region","Site.name", "Subregion","Site"))%>% filter(!is.na(ndvi_ED50)) 

dat_all_touse_fvfmrel9c <- dat_all_touse%>% dplyr::select(-c("fvfm_ED50","ndvi_ED50","RelativeNDVI9C", "fvfm_ED50CIrange","ndvi_ED50CIrange","Site_geno_year","Region","Site.name", "Subregion","Site")) %>% filter(!is.na(RelativeFvFm9C)) 

dat_all_touse_ndvirel9c <- dat_all_touse  %>% dplyr::select(-c("fvfm_ED50","ndvi_ED50",  "RelativeFvFm9C","fvfm_ED50CIrange","ndvi_ED50CIrange","Site_geno_year","Region","Site.name", "Subregion","Site"))%>% filter(!is.na(RelativeNDVI9C)) 


cols = c("Temperature: Climatology" = "royalblue2", "Temperature: Recent Heat Stress" = "navyblue", "Depth" = "lightsteelblue",  "Temperature: Anomalies" = "deepskyblue2","Symbiont Community" = "tomato3" , "Latitude" = "seagreen4", "Temperature: Variability" = "cornflowerblue", "Nutrient Availability" = "orange", "Pigmentation" = "violetred3", "Host Genomic Cluster" = "darkgoldenrod2", "Latitude" = "red")


```

#Make a correlation matrix of predictors going into BRTs
Some variables are correlated 
```{r}

library(Hmisc)
library(corrplot)
library(factoextra)
library(reshape2)

get_upper_tri <- function(cormat){
  cormat[lower.tri(cormat)]<- NA
  return(cormat)}
reorder_cormat <- function(cormat){
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]}
env.data_corr <- dat_all_touse %>% dplyr::select(-c("fvfm_ED50","ndvi_ED50",  "RelativeFvFm9C","RelativeNDVI9C","fvfm_ED50CIrange","ndvi_ED50CIrange","Site_geno_year","Region","Site.name", "Subregion","Site","dominant_type.adj","Clust","Shelf.pos"))
rcorr(as.matrix(env.data_corr, type = c("pearson","spearman")))
cor <- cor(env.data_corr, use="complete.obs")
cormat <- reorder_cormat(cor)
upper_tri <- get_upper_tri(cormat)
melted_cormat <- melt(upper_tri, na.rm = TRUE)
ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 90, vjust = 1, 
    size = 10, hjust = 1), axis.text.y = element_text(size = 10))+
 coord_fixed() + xlab("") + ylab("")
ggsave(paste0("../Outputs/BRTs/corrmatrix_brt",date,".png"), width = 8, height = 7)

```

##BRT for FvFm ed50
Workflow:
-use gbm.step to get total and cv deviance to get deviance explained; and to get # of trees 
-lr, tree complexity, and bag fraction set to be conservative and allow slow learning rate 
-additional parameter search was completed beforehand without much change in results using diff lr, tc, and bf values 
```{r }

set.seed(123)
brt.fvfmed50 <- gbm.step(data = dat_all_fvfmED50_filt,
  gbm.x = c(1:26,28),
  gbm.y = 27,
  family = "gaussian",
  tree.complexity = 5,
  learning.rate = 0.001,
  bag.fraction = 0.5)  
```

```{r}
# mean total deviance = 0.436 
# mean residual deviance = 0.095 
# estimated cv deviance = 0.144 

Dfvfmed50 = (0.436 - 0.144) / 0.436

gbm.plot(brt.fvfmed50)

#use number of trees suggested above 
brt.fvfmed50.gbm <- gbm(formula = fvfm_ED50 ~ .,
  distribution = "gaussian",
  data = dat_all_fvfmED50_filt,
  n.trees = 4800,
  interaction.depth = 5,
  shrinkage = 0.001,
  bag.fraction = 0.5,
  cv.folds = 10)
summary(brt.fvfmed50.gbm)
```
#Plot top variables and partial dependance plots 
```{r }
brt.fvfm.best_pred <- summary(brt.fvfmed50, cBars = 10,method = relative.influence,  las = 2)%>% left_join(pred_types, by = "var")

p1 <- brt.fvfm.best_pred %>% slice(1:4) %>% ggplot(aes(reorder(name, rel.inf), rel.inf, fill = Predictor.Type)) + geom_bar(stat = "identity") + theme_bw() + coord_flip() + ylab("Relative Influence") + xlab("")  + ggtitle("ED50 (Fv/Fm) \nDeviance explained: 0.670")+ ylim(0,42)+ scale_fill_manual(name = "Predictor Type", values = cols)

MMM <- pdp::partial(brt.fvfmed50.gbm, pred.var = "MMM", n.trees = 4800)%>% autoplot() + theme_bw() + xlab("Maximum Monthly Mean (˚C)") + ylab("ED50 (Fv/Fm)") + ylim(c(36.15,36.65))
 
dhw_coll <- pdp::partial(brt.fvfmed50.gbm, pred.var = "DHW_collection", n.trees = 4800)%>% autoplot() + theme_bw() + xlab("DHW at time of collection") + ylab("ED50 (Fv/Fm)")+ ylim(c(36.15,36.65))

pfvfmed50 <- p1 + MMM + dhw_coll + plot_layout(guides = 'collect')+ plot_annotation(tag_levels = 'A') 

#similar between two slightly diff brt functions 
summary(brt.fvfmed50)
summary(brt.fvfmed50.gbm)


```

#ndvi ed50 
```{r ndvi_ed50}
set.seed(123)

brt.ndvied50 <- gbm.step(data = dat_all_ndviED50_filt,
  gbm.x = c(1:26,28),
  gbm.y = 27,
  family = "gaussian",
  tree.complexity = 5,
  learning.rate = 0.001,
  bag.fraction = 0.5)  
```

#rerun with gbm and correct ideal number of trees
```{r }
# mean total deviance = 0.883 
# mean residual deviance = 0.436 
# estimated cv deviance = 0.566


(0.883 - 0.566) / 0.883

summary(brt.ndvied50)
gbm.plot(brt.ndvied50)

brt.ndvied50.gbm <- gbm(formula = ndvi_ED50 ~ .,
  distribution = "gaussian",
  data = dat_all_ndviED50_filt,
  n.trees = 2350,
  interaction.depth = 5,
  shrinkage = 0.001,
  bag.fraction = 0.5,
  cv.folds = 10)
summary(brt.ndvied50.gbm)
```

#variable and pdp plots 
```{r }

MMM2 <- pdp::partial(brt.ndvied50.gbm, pred.var = "MMM", n.trees = 2350)%>% autoplot() + theme_bw() + xlab("Maximum Monthly Mean (˚C)") + ylab("ED50 (NDVI)") +ylim(c(35.35,36))

om <- pdp::partial(brt.ndvied50.gbm, pred.var = "OM", n.trees = 2350)%>% autoplot() + theme_bw() + xlab("Mean Temperature (˚C)") + ylab("ED50 (NDVI)")+ylim(c(35.35,36))

brt.ndvied50_pred <- summary(brt.ndvied50, cBars = 10,method = relative.influence,  las = 2)%>% left_join(pred_types, by = "var")
p1 <- brt.ndvied50_pred %>% slice(1:4) %>% ggplot(aes(reorder(name, rel.inf), rel.inf, fill = Predictor.Type)) + geom_bar(stat = "identity") + theme_bw() + coord_flip() + ylab("Relative Influence") + xlab("")  + ggtitle("ED50 (NDVI) \nDeviance explained: 0.359")+ ylim(0,42)+ scale_fill_manual(name = "Predictor Type", values = cols)

pndvied50 <- p1 + MMM2 + om + plot_layout(guides = 'collect')+ plot_annotation(tag_levels = 'A') 

summary(brt.ndvied50)
summary(brt.ndvied50.gbm)


```

#retained fvfm 
```{r}
set.seed(123)
brt.fvfmrel <- gbm.step(gbm.x = c(1:26,28), gbm.y = 27,
  family = "gaussian",
  data = dat_all_touse_fvfmrel9c,
  tree.complexity = 5,
  learning.rate = 0.001,
  bag.fraction = 0.5)  
```


```{r }
# mean total deviance = 0.031 
# mean residual deviance = 0.004 
# estimated cv deviance = 0.006 

(0.031 - 0.006)/0.031

set.seed(123)
brt.fvfmrel.gbm <- gbm(formula = RelativeFvFm9C ~ .,
  distribution = "gaussian",
  data = dat_all_touse_fvfmrel9c,
  n.trees = 4950,
  interaction.depth = 5,
  shrinkage = 0.001,
  bag.fraction = 0.5,
  cv.folds = 10)

summary(brt.fvfmrel)
summary(brt.fvfmrel.gbm)

lat <- pdp::partial(brt.fvfmrel.gbm, pred.var = "Lat", n.trees = 4950)%>% autoplot() + theme_bw() + xlab("Latitude") + ylab("Retained FvFm")


dhw8 <- pdp::partial(brt.fvfmrel.gbm, pred.var = "DHW_freq_sup8", n.trees = 4950)%>% autoplot() + theme_bw() + xlab("Frequency of DHW > 8") + ylab("Retained FvFm")


brt.fvfmrel_pred <- summary(brt.fvfmrel, cBars = 10,method = relative.influence,  las = 2)%>% left_join(pred_types, by = "var")
p1 <- brt.fvfmrel_pred %>% slice(1:4) %>% ggplot(aes(reorder(name, rel.inf), rel.inf, fill = Predictor.Type)) + geom_bar(stat = "identity") + theme_bw() + coord_flip() + ylab("Relative Influence") + xlab("")  + ggtitle("Retained FvFm \nDeviance explained: 0.806")+ ylim(0,42)+ scale_fill_manual(name = "Predictor Type", values = cols)


pfvfmrel <- p1 + lat + dhw8 + plot_layout(guides = 'collect')+ plot_annotation(tag_levels = 'A') 


```

##relative ndvi
```{r}

brt.ndvirel <- gbm.step(gbm.x = c(1:26,28), gbm.y = 27,
  family = "gaussian",
  data = dat_all_touse_ndvirel9c,
  tree.complexity = 5,
  learning.rate = 0.001,
  bag.fraction = 0.5)  
```

```{r }
# mean total deviance = 0.025 
# mean residual deviance = 0.009 
# estimated cv deviance = 0.013 

(0.025 - 0.013)/ 0.025


set.seed(123)
brt.ndvirel.gbm <- gbm(formula = RelativeNDVI9C ~ .,
  distribution = "gaussian",
  data = dat_all_touse_ndvirel9c,
  n.trees = 3950,
  interaction.depth = 5,
  shrinkage = 0.001,
  bag.fraction = 0.5,
  cv.folds = 10)
```

```{r }

DHW_collection <- pdp::partial(brt.fvfmrel.gbm, pred.var = "DHW_collection", n.trees = 3950)%>% autoplot() + theme_bw() + xlab("DHW at time of collection") + ylab("Retained NDVI")+ ylim(c(0.33,0.44))

TSA_DHW_mean <- pdp::partial(brt.fvfmrel.gbm, pred.var = "TSA_DHW_mean", n.trees = 3950)%>% autoplot() + theme_bw() + xlab("Mean DHW") + ylab("Retained NDVI")+ ylim(c(0.33,0.44))

brt.ndvirel_pred <- summary(brt.ndvirel, cBars = 10,method = relative.influence,  las = 2)%>% left_join(pred_types, by = "var")
p1 <- brt.ndvirel_pred %>% slice(1:4) %>% ggplot(aes(reorder(name, rel.inf), rel.inf, fill = Predictor.Type)) + geom_bar(stat = "identity") + theme_bw() + coord_flip() + ylab("Relative Influence") + xlab("")  + ggtitle("Retained NDVI \nDeviance explained: 0.480")+ ylim(0,42)+ scale_fill_manual(name = "Predictor Type", values = cols)


pndvirel <- p1 + DHW_collection + TSA_DHW_mean + plot_layout(guides = 'collect')+ plot_annotation(tag_levels = 'A') 


```


#visualize as one figure
##Figure 6
```{r}

((pfvfmed50) / (pfvfmrel) / (pndvied50) / (pndvirel)) + plot_layout(guides = 'collect') & theme(axis.text.y = element_text(size = 14), legend.text = element_text(size = 14), legend.title = element_text(size = 14), axis.title.x = element_text(size = 14), axis.title.y = element_text(size = 14), title = element_text(size = 14.5))
ggsave(paste0("../Outputs/BRTs/Figure6_",date,".png"), width = 16, height = 10)

```








