---
title: "Data_overview"
author: "Melissa Naugle"
date: "1/12/2023"
output: html_document
---

# Exloratory script of heat tolerance data
- Calculates summary statistics (for Table 1)
- Plots heat tolerance performance among colonies and sites (Fig 2)
- Explores distributions and correlations among phenotypic traits (Fig S2, S3)
- Measures relationship between NDVI and lab-assayed chlorophyll (Fig S1)
- Generates sampling figure (Fig 1)
- Assess variation within versus among sites (Table 1)

```{r setup, include=FALSE}
setwd("/Users/Melissa/Desktop/GitHub/Ahya_heat_tol_variation//Rscripts_analysis/")
rm(list = ls())
graphics.off()
library(tidyverse)
library(patchwork)
library(tidytext)
library(PupillometryR)
date = format(Sys.Date(), "%Y_%m_%d")
outputpath <- "/Users/Melissa/Desktop/GitHub/Ahya_heat_tol_variation/Outputs/Data_exploration/"
```

#data in


```{r}
pheno <- read.csv("../CleanData/pheno_metadata_clean_2024_01_08.csv") 
lat <- pheno %>% dplyr::select(Site.name,Lat) %>% distinct()
env <- read.csv("../CleanData/Environmental_data_PCs_2024_01_08.csv")
mmm <- env %>% dplyr::select(Site.name, MMM)%>% distinct()
coast <- pheno %>% dplyr::select(Site.name, Km_to_coastline) %>% distinct()
clust <- read.csv("../CleanData/2023_10_17_clust_data_edited.csv")
symbiont <- read.csv("../CleanData/ITS2_types_OTUs_17_10_2023.csv")
chlorophyll <- read.csv(("../Rawdata/TotalChlorophyll.csv")) %>% 
  dplyr::select(-Species) %>% mutate(Genotype = as.character(Genotype), Rack.Pos = as.character(Rack.Pos))
rawfvfm <- read.csv("../Rawdata/full_iPAM_data_2023_12_04.csv") 
rawfvfm$Rack.Pos <-  sapply(strsplit(as.character(rawfvfm$Site_Rack), "_"), `[`, 2)
rawndvi <- read.csv("../Rawdata/hyperspecdata_2023_12_04.csv") %>% mutate(Genotype = as.character(Genotype), Rack.Pos = as.character(Rack.Pos))

```




#summary stats
Mean, median, and range for each trait
```{r}
print("fvfm_ED50")
mean(pheno$fvfm_ED50, na.rm = T)
range(pheno$fvfm_ED50, na.rm = T)
median(pheno$fvfm_ED50, na.rm = T)

print("RelativeFvFm9C")
mean(pheno$RelativeFvFm9C, na.rm = T)
range(pheno$RelativeFvFm9C, na.rm = T)
median(pheno$RelativeFvFm9C, na.rm = T)

print("ndvi_ED50")
mean(pheno$ndvi_ED50, na.rm = T)
range(pheno$ndvi_ED50, na.rm = T)
median(pheno$ndvi_ED50, na.rm = T)

print("RelativeNDVI9C")
mean(pheno$RelativeNDVI9C, na.rm = T)
range(pheno$RelativeNDVI9C, na.rm = T)
median(pheno$RelativeNDVI9C, na.rm = T)
```


#Explore distributions of the 4 traits/metrics
```{r}

#by subregion
pheno$Subregion <- factor(pheno$Subregion, levels = c("Cooktown", "Cairns", "Innisfail/Townsville", "Townsville" , "Swains", "Capricorn Bunker"))
cols <-  c("Cooktown" = "red2", "Cairns" = "sienna1", "Innisfail/Townsville" = "gold2", "Townsville" = "seagreen3", "Swains" = "cornflower blue", "Capricorn Bunker" = "blue")

p1 <- pheno %>% ggplot(aes(fvfm_ED50)) + geom_histogram(aes(fill = Subregion)) + theme_bw() + scale_color_manual(values = cols) + theme(legend.position = "none")
p2 <- pheno %>% ggplot(aes(RelativeFvFm9C)) + geom_histogram(aes(fill = Subregion)) + theme_bw() + theme(legend.position = "none") + ylab("")+ scale_color_manual(values = cols)

p4 <- pheno %>% ggplot(aes(ndvi_ED50)) + geom_histogram(aes(fill = Subregion)) + theme_bw() + theme(legend.position = "none")+ scale_color_manual(values = cols)
p5 <- pheno %>% ggplot(aes(RelativeNDVI9C)) + geom_histogram(aes(fill = Subregion)) + theme_bw() + theme(legend.position = "none")+ ylab("")+ scale_color_manual(values = cols)

(p1+p2) / (p4 + p5)+ plot_layout(guides = "collect") & theme(legend.position = "top")
```

#is there a relationship between traits?
##Figure S3
```{r }
p1 <- pheno %>% ggplot(aes(RelativeFvFm9C, RelativeNDVI9C)) + geom_point() + theme_bw()+ geom_smooth(method = "lm", se = F, color = "blue")  + ggpubr::stat_cor(method = "pearson", size = 4)+ xlab("Retained Fv/Fm") + ylab("Retained NDVI")

p3 <- pheno %>% ggplot(aes(fvfm_ED50, ndvi_ED50)) + geom_point() + theme_bw()+ geom_smooth(method = "lm", se = F, color = "blue")  + ggpubr::stat_cor(method = "pearson", size = 4)+ xlab("ED50 Fv/Fm") + ylab("ED50 NDVI")

p3 + p1 
ggsave(paste0(outputpath,"FigureS3_",date,".png"), width = 7, height = 4)
```



#is there a relationship between the metrics computed for each trait?
##Figure S2
```{r }
p1 <- pheno %>% ggplot(aes(RelativeFvFm9C, fvfm_ED50)) + geom_point() + theme_bw()+ geom_smooth(method = "lm", se = F, color = "blue")  + ggpubr::stat_cor(method = "pearson", size = 4) + xlab("Relative Fv/Fm") + ylab("ED50 Fv/Fm") + xlim(c(0,0.8))

p2 <- pheno %>% ggplot(aes(RelativeNDVI9C, ndvi_ED50)) + geom_point() + theme_bw()+ geom_smooth(method = "lm", se = F, color = "blue")  + ggpubr::stat_cor(method = "pearson", size = 4)+ xlab("Relative NDVI") + ylab("ED50 NDVI")+ xlim(c(0,0.8))

p1 + p2
#weak relationship 

p1 <- pheno %>% ggplot(aes(RelativeFvFm9C, fvfm_ED50)) + geom_point() + theme_bw()+ geom_smooth(method = "lm", se = F, color = "blue")  + ggpubr::stat_cor(method = "pearson", size = 4) + xlab("Relative Fv/Fm") + ylab("ED50 Fv/Fm") + xlim(c(0,0.8)) + ylim(c(33,40)) + facet_wrap(~Site.name) 

p2 <- pheno %>% ggplot(aes(RelativeNDVI9C, ndvi_ED50)) + geom_point() + theme_bw()+ geom_smooth(method = "lm", se = F, color = "blue")  + ggpubr::stat_cor(method = "pearson", size = 4)+ xlab("Relative NDVI") + ylab("ED50 NDVI")+ xlim(c(0,0.8)) + ylim(c(33,40))+ facet_wrap(~Site.name)

p1 + p2
#relationship exists within sites though 


p1 <- pheno %>% ggplot(aes(RelativeFvFm9C.adj, fvfm_ED50.adj)) + geom_point() + theme_bw()+ geom_smooth(method = "lm", se = F, color = "blue")  + ggpubr::stat_cor(method = "pearson", size = 4) + xlab("Retained Fv/Fm (site-adjusted)") + ylab("ED50 Fv/Fm (site-adjusted)") + xlim(c(0,0.65))+ ylim(c(35,37.5))

p2 <- pheno %>% ggplot(aes(RelativeNDVI9C.adj, ndvi_ED50.adj)) + geom_point() + theme_bw()+ geom_smooth(method = "lm", se = F, color = "blue")  + ggpubr::stat_cor(method = "pearson", size = 4)+ xlab("Retained NDVI (site-adjusted)") + ylab("ED50 NDVI (site-adjusted)")+ xlim(c(0,0.75))+ ylim(c(35,38.5))

p1 + p2
ggsave(paste0(outputpath,"FigureS2_",date,".png"), width = 7, height = 4)


```

#top 20% of performers according to each metric 

```{r}
pheno2_top.performers <- pheno %>% 
    arrange(-fvfm_ED50) %>% mutate(Ranking_fvfm_ED50 = row_number()) %>% 
    mutate(Ranking_fvfm_ED50 = replace(Ranking_fvfm_ED50, Ranking_fvfm_ED50 < (nrow(pheno)*0.2), "Top 20")) %>% 
    mutate(Ranking_fvfm_ED50 = replace(Ranking_fvfm_ED50, Ranking_fvfm_ED50 != "Top 20", "Bottom 80")) %>%
    arrange(-ndvi_ED50) %>% mutate(Ranking_ndvi_ED50 = row_number()) %>% 
    mutate(Ranking_ndvi_ED50 = replace(Ranking_ndvi_ED50, Ranking_ndvi_ED50 < (nrow(pheno)*0.2), "Top 20")) %>% 
    mutate(Ranking_ndvi_ED50 = replace(Ranking_ndvi_ED50, Ranking_ndvi_ED50 != "Top 20", "Bottom 80")) %>%
    arrange(-RelativeFvFm9C) %>% mutate(Ranking_fvfm_rel9c = row_number()) %>%
    mutate(Ranking_fvfm_rel9c = replace(Ranking_fvfm_rel9c, Ranking_fvfm_rel9c < (nrow(pheno)*0.2), "Top 20"))%>% 
    mutate(Ranking_fvfm_rel9c = replace(Ranking_fvfm_rel9c, Ranking_fvfm_rel9c != "Top 20", "Bottom 80")) %>%
     arrange(-RelativeNDVI9C) %>% mutate(Ranking_ndvi_rel9c = row_number()) %>%
    mutate(Ranking_ndvi_rel9c = replace(Ranking_ndvi_rel9c, Ranking_ndvi_rel9c < (nrow(pheno)*0.2), "Top 20"))%>% 
    mutate(Ranking_ndvi_rel9c = replace(Ranking_ndvi_rel9c, Ranking_ndvi_rel9c != "Top 20", "Bottom 80")) %>%
    dplyr::select(Site.name, Ranking_fvfm_ED50,Ranking_ndvi_ED50,Ranking_fvfm_rel9c,Ranking_ndvi_rel9c, Site_year, Genotype, Lat, Lon)


pheno2_top.performers$top20s <- rowSums(pheno2_top.performers == "Top 20")

pheno2_top.performers$fvfm[pheno2_top.performers$Ranking_fvfm_ED50 == "Top 20" | pheno2_top.performers$Ranking_fvfm_rel9c == "Top 20"] <- 1
pheno2_top.performers$fvfm[pheno2_top.performers$Ranking_fvfm_ED50 == "Top 20" & pheno2_top.performers$Ranking_fvfm_rel9c == "Top 20"] <- 2

pheno2_top.performers$ndvi[pheno2_top.performers$Ranking_ndvi_ED50 == "Top 20" | pheno2_top.performers$Ranking_ndvi_rel9c == "Top 20"] <- 1
pheno2_top.performers$ndvi[pheno2_top.performers$Ranking_ndvi_ED50 == "Top 20" & pheno2_top.performers$Ranking_ndvi_rel9c == "Top 20"] <- 2

pheno2_top.performers$ed50[pheno2_top.performers$Ranking_ndvi_ED50 == "Top 20" | pheno2_top.performers$Ranking_fvfm_ED50 == "Top 20"] <- 1
pheno2_top.performers$ed50[pheno2_top.performers$Ranking_ndvi_ED50 == "Top 20" & pheno2_top.performers$Ranking_fvfm_ED50 == "Top 20"] <- 2

pheno2_top.performers$rel[pheno2_top.performers$Ranking_fvfm_rel9c == "Top 20" | pheno2_top.performers$Ranking_ndvi_rel9c == "Top 20"] <- 1
pheno2_top.performers$rel[pheno2_top.performers$Ranking_fvfm_rel9c == "Top 20" & pheno2_top.performers$Ranking_ndvi_rel9c == "Top 20"] <- 2

pheno2_top.performers


pheno2_top.performers %>% filter(fvfm == 2)
pheno2_top.performers %>% filter(ndvi == 2)
pheno2_top.performers %>% filter(ed50 == 2)
pheno2_top.performers %>% filter(rel == 2)
```


#Which colonies were within the top 20% of performers according to each trait/metric? 
#Where do these top performing colonies occur (north/south and which sites)?
##Figure 2
```{r}
pheno2_top.performers_2 <- pheno2_top.performers %>% left_join(pheno, by = c("Site_year", "Genotype", "Site.name", "Lat","Lon"))

pheno2_top.performers_2$Ranking_fvfm_ED50[pheno2_top.performers_2$Ranking_fvfm_ED50 == "Top 20"] <- "Top 20 (Absolute)"
pheno2_top.performers_2$Ranking_ndvi_ED50[pheno2_top.performers_2$Ranking_ndvi_ED50 == "Top 20"] <- "Top 20 (Absolute)"
pheno2_top.performers_2$Ranking_fvfm_rel9c[pheno2_top.performers_2$Ranking_fvfm_rel9c == "Top 20"] <- "Top 20 (Retained)"
pheno2_top.performers_2$Ranking_ndvi_rel9c[pheno2_top.performers_2$Ranking_ndvi_rel9c == "Top 20"] <- "Top 20 (Retained)"

p1 <- pheno2_top.performers_2 %>%
  filter(!is.na(fvfm_ED50)) %>%
  mutate(Site.name = fct_reorder(Site.name, Lat)) %>%
  ggplot(aes(x = reorder(Site.name, Lat), y = fvfm_ED50, group = Site.name)) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
                   adjust = 1, width = 2, alpha = 0.7, lwd = 0) +
  geom_boxplot(position = position_nudge(x = -0.2, y = 0), width = 0.1,
               outlier.shape = NA, lwd = 0.25) +
  geom_point(aes(color = Ranking_fvfm_ED50), alpha = 0.6) +
  coord_flip() +
  theme_bw() +
  labs(y = "Absolute ED50 (Fv/Fm)", x = "") + 
  scale_color_manual(name = "Ranking", values = c("darkgrey","red"))


p2 <- pheno2_top.performers_2 %>%
  filter(!is.na(ndvi_ED50)) %>%
  mutate(Site.name = fct_reorder(Site.name, Lat)) %>%
  ggplot(aes(x = reorder(Site.name, Lat), y = ndvi_ED50, group = Site.name)) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
                   adjust = 1, width = 2, alpha = 0.7, lwd = 0) +
  geom_boxplot(position = position_nudge(x = -0.2, y = 0), width = 0.1,
               outlier.shape = NA, lwd = 0.25) +
  geom_point(aes(color = Ranking_ndvi_ED50), alpha = 0.6) +
  coord_flip() +
  theme_bw() +
  labs(y = "Absolute ED50 (NDVI)", x = "") + 
  scale_color_manual(name = "Ranking", values = c("darkgrey","red"))+ 
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank())


p3 <- pheno2_top.performers_2 %>%
  filter(!is.na(RelativeFvFm9C)) %>%
  mutate(Site.name = fct_reorder(Site.name, Lat)) %>%
  ggplot(aes(x = reorder(Site.name, Lat), y = RelativeFvFm9C, group = Site.name)) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
                   adjust = 1, width = 2, alpha = 0.7, lwd = 0) +
  geom_boxplot(position = position_nudge(x = -0.2, y = 0), width = 0.1,
               outlier.shape = NA, lwd = 0.25) +
  geom_point(aes(color = Ranking_fvfm_rel9c), alpha = 0.6) +
  coord_flip() +
  theme_bw() +
  labs(y = "Retained Fv/Fm", x = "") + 
  scale_color_manual(name = "Ranking", values = c("darkgrey","blue"))


p4 <- pheno2_top.performers_2 %>%
  filter(!is.na(RelativeNDVI9C)) %>% 
   add_row(Site.name = "Reef 21550", Lat = -21.9455) %>%
  mutate(Site.name = fct_reorder(Site.name, Lat)) %>%
  ggplot(aes(x = reorder(Site.name, Lat), y = RelativeNDVI9C, group = Site.name)) +
  geom_flat_violin(position = position_nudge(x = 0.2, y = 0), 
                   adjust = 1, width = 2, alpha = 0.7, lwd = 0) +
  geom_boxplot(position = position_nudge(x = -0.2, y = 0), width = 0.2,
               outlier.shape = NA, lwd = 0.25) +
  geom_point(aes(color = Ranking_ndvi_rel9c), alpha = 0.6) +
  coord_flip() +
  theme_bw() +
  labs(y = "Retained NDVI", x = "") + 
  scale_color_manual(name = "Ranking", values = c("darkgrey","blue")) + 
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) 


pLAT <- pheno2_top.performers_2 %>% filter(!Site.name == "Reef21550") %>% select(Site.name,Lat) %>% distinct %>% ggplot(aes(x = reorder(Site.name, Lat),  fill = Lat)) + geom_bar(position = "fill", width = 1) + coord_flip()+ ylab("") + xlab("") + theme_minimal() + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) + scale_fill_continuous(name = "Latitude")+ 
  theme(axis.text.y = element_blank())


pLAT + p1 + plot_layout(ncol = 2, widths = c(1,20), guides = "collect")
(p1 + p2) / (p3 + p4) + plot_layout(guides = "collect")+ plot_annotation(tag_levels = 'A') & 
  theme(text = element_text(size = 14))

ggsave(paste0(outputpath,"Figure2_",date,".png"), width =8, height = 7)


```


#How does image-based NDVI relate to lab-assayed chlorophyll?
##Figure S1

```{r}
rawfvfm <- rawfvfm %>% mutate(Genotype = as.character(Genotype))
pheno_chlorophyll_fragment <-  left_join(chlorophyll, rawndvi, by = c("Site", "Genotype", "SampleYear", "Rack.Pos", "Treatment")) %>% left_join(rawfvfm, by = c("Site", "Genotype", "SampleYear", "Rack.Pos"), suffix = c("", ".y")) %>% dplyr::select(-ends_with(".y"))

pheno_chlorophyll_colony <- pheno_chlorophyll_fragment  %>% dplyr::group_by(Site, SampleYear, Genotype, Treatment) %>% dplyr::summarise( MedNDVIgeno = mean(MedNDVI), chltot_dryweight = mean(Total.Chl..ug.mg.dry.weight.)) 

formula <- y ~ poly(x, 3, raw = TRUE)


a <- pheno_chlorophyll_fragment %>% ggplot(aes(MedNDVI, `Total.Chl..ug.mg.dry.weight.`)) + geom_point(aes(col = Site, shape = Treatment)) + theme_bw() + geom_smooth(method = "lm", se = F, color = "darkgrey", formula = y ~ x + I(x^2), size = 1) +  theme(legend.position = "none", plot.title = element_text(hjust = 0.5))+ ylab("Total Chlorophyll / dry weight") + xlab("Median NDVI") + ggtitle("Fragment-level") + ggpubr::stat_cor(method = "spearman", size = 4.9)+ theme( plot.title = element_text(hjust = 0.5), text = element_text(size = 16), title = element_text(size = 14)) + ylim(c(0,0.1))

b <- pheno_chlorophyll_colony  %>% ggplot(aes(MedNDVIgeno, chltot_dryweight)) + geom_point(aes(col = Site, shape = Treatment)) + theme_bw() + geom_smooth(method = "lm", se = F, color = "darkgrey", formula = y ~ x + I(x^2), size = 1) + ylab("")+ xlab("Median NDVI") + ggpubr::stat_cor(method = "spearman", size = 4.9)+ ggtitle("Colony-level") + theme( plot.title = element_text(hjust = 0.5), text = element_text(size = 16), title = element_text(size = 14))+ ylim(c(0,0.1))


a + b+ plot_annotation(tag_levels = 'A') 

ggsave(paste0(outputpath,"FigureS1_",date,".png"), width = 9, height = 4)

b

#ggsave(paste0(outputpath,"NDVI_chlorophyll_colonylevel",date,".png"), width = 6, height = 4)

```




#map of sample reefs 
##Figure 1

```{r}

library("rnaturalearthdata")
library("rnaturalearth")
library("sf")
library(packcircles)
library(tmap)

sites <- pheno %>% dplyr::select(Lon, Lat, Site.name, Shelf.pos) %>% left_join(mmm, by = "Site.name") %>% group_by(Site.name, Lon, Lat, Shelf.pos,MMM) %>% tally() %>% dplyr::rename('Sample Size' = n) %>% distinct() %>% filter(!MMM == "28.591") #remove hicks_2

sites$Radius <- 0.3

coords_before <- sites %>% dplyr::select(Lon, Lat, Radius, Site.name, Shelf.pos, MMM)
coords_after <- as.data.frame(packcircles::circleLayout(coords_before))
coords_after$Site.name <- sites$Site.name
coords_after$Shelf.pos <- sites$Shelf.pos
coords_after$`Sample Size` <- sites$`Sample Size`
coords_after$MMM <- sites$MMM
coords_after <-  coords_after %>% dplyr::select(layout.x, layout.y, Site.name, Shelf.pos, `Sample Size`, MMM)


world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)

pmap <- ggplot(data = world) + 
  geom_sf(fill = "gray70") + 
  coord_sf(xlim = c(141, 155), ylim = c(-26, -11), expand = FALSE) + 
  theme_bw()  + xlab("") + ylab("")+ 
  geom_point(data = coords_after, aes(layout.x, layout.y, fill = MMM, size = `Sample Size`), alpha = 0.8, pch = 21) +scale_fill_viridis_c() + ggrepel::geom_text_repel(data = coords_after, aes(layout.x, layout.y, label = Site.name), size = 5) + theme(legend.position = "top") + scale_radius(range = c(6,10)) + 
  guides(fill = guide_colourbar(title.position="top", title.hjust = 0.5),
         size = guide_legend(title.position="top", title.hjust = 0.5))



gbr_reefs = rgdal::readOGR("../Rawdata/GBR_features2/Marine_Bioregions_of_the_Great_Barrier_Reef__Reef_.shp")
AG <- fortify(gbr_reefs)

pmap2 <- ggplot(data = world) + 
  geom_sf(fill = "gray70") + 
  coord_sf(xlim = c(141, 155), ylim = c(-26, -11), expand = FALSE) + geom_polygon(data=AG, aes(long, lat, group = group), colour = alpha("grey25", 1/2), size = 0.5)+ 
  theme_bw()  + xlab("") + ylab("")+ 
  geom_point(data = sites, aes(Lon, Lat, fill = MMM), alpha = 0.8, pch = 21, size = 3) +scale_fill_viridis_c(limits = c(27,29)) + ggrepel::geom_label_repel(data = sites, aes(Lon, Lat, label = Site.name, fill = MMM), size = 5.5, col = "white") + theme(legend.position = c(.9,.88), text = element_text(size = 14)) 

pmap2

ggsave(paste0(outputpath,"Figure1_map_reefs_",date,".png"), width = 7, height = 8)

```





#how much variation exists within vs among sites?
#how much variation exists within vs among sites and putative species?
##Table 1

```{r}
library(lme4)
m1 <- lmer(fvfm_ED50 ~ (1|Site.name), data = pheno)
summary(m1)
0.3035/0.1900 #1.60
mean(pheno$fvfm_ED50, na.rm = T)
range(pheno$fvfm_ED50, na.rm = T)

m1 <- lmer(RelativeFvFm9C ~ (1|Site.name), data = pheno)
summary(m1)
0.044370/0.007378 #6.01
mean(pheno$RelativeFvFm9C, na.rm = T)
range(pheno$RelativeFvFm9C, na.rm = T)

m1 <- lmer(ndvi_ED50 ~ (1|Site.name), data = pheno)
summary(m1)
0.5234/0.3578 #1.46
mean(pheno$ndvi_ED50, na.rm = T)
range(pheno$ndvi_ED50, na.rm = T)

m1 <- lmer(RelativeNDVI9C ~ (1|Site.name), data = pheno)
summary(m1)
0.01361/0.01052 #1.29
mean(pheno$RelativeNDVI9C, na.rm = T)
range(pheno$RelativeNDVI9C, na.rm = T)

pheno_clust <- left_join(pheno, clust, by = c("Sitename_geno_year", "Site.name"))

#account for putative species
m1 <- lmer(fvfm_ED50 ~ (1|Site.name) + (1|Clust), data = pheno_clust)
summary(m1)
0.301729/0.188341 #1.60
m2 <- lmer(RelativeFvFm9C ~ (1|Site.name)+ (1|Clust), data = pheno_clust)
summary(m2) 
0.0434084/0.0072400 #6.00
m3 <- lmer(ndvi_ED50 ~ (1|Site.name)+ (1|Clust), data = pheno_clust)
summary(m3)
0.51893/0.34939 #1.49
m4 <- lmer(RelativeNDVI9C ~ (1|Site.name)+ (1|Clust), data = pheno_clust)
summary(m4)
0.01364/0.01051 #1.30
pheno_clust %>% filter(!is.na(RelativeNDVI9C)) %>% group_by(Site.name, Clust) %>% tally()

```




